Template Master's Thesis

### 2023-08-07 paper.tex
Update paper

%However, they are not widely used even cannot guarantee perfect privacy (factors: size of anonymity set, usage /user behaviour of protocol, privacy guarantees depend on other users).
%The most prominent/widely used non-custodial mixer is Tornado cash (Schär, Tutela).
%This reaction has, in turn, sparked the development of other clustering heuristics exploiting user behavior interacting with privacy-enhancing protocols.

\textbf{Related work}\\
Previous work in this field can be broadly distinguished between heuristics developed for the unspent transaction output (UTXO) model (e.g Bitcoin, ZCash) and the account-based model (e.g. Ethereum, Polygon). While both models share the concept of addresses, the notion of accounts is not present in UTXO-based blockchains. The way in which transactions are processed is fundamentally different, and therefore, clustering heuristics are not applicable to both paradigms.
%\textbf{UTXO-model clustering heuristics} 
A number of address clustering heuristics in the UTXO setup have been proposed for Bitcoin and derivatives \citep{Androulaki2013, Meiklejohn2013, Haslhofer2016, jourdan2018, kappos2022}. All these methods will not be the focus of this work and therefore not discussed. As suggested by Nakamoto in the Bitcoin whitepaper, most Bitcoin wallet implementations use a new key pair for each transaction to keep them from being linked to a common owner. \newline
In contrast to the UTXO-model, native transactions in account-based blockchains can only move funds between a single sender and receiver, and the ``change``remains in the sender account. Subsequent transactions necessarily use the same address again. The account-based model essentially relies on address reuse on the protocol level. Therefore, privacy guarantees should be lower and most users likely use only a small number of addresses. \citep{Beres2020} \newline
% Make an assumption that people only use a small set of addressesyy
%\textbf{Account-model clustering heuristics}
Clustering heuristics for account-based blockchains were first introduced by \cite{FV:17}. He proposed and applied heuristics that exploit patterns related to deposit addresses, multiple participation in airdrops, and token authorization mechanisms. \newline
\cite{Beres2020} propose more universally applicable methods, as they argue that Victor’s heuristics, while powerful, assume participation in certain on-chain events. Their technique interprets transactions or token transfers as network graphs, with addresses as nodes and asset transfers as edges. They quantitavely compare graph-representation learning algorithms (a subset of machine learning) and propose further user profiling techniques based on time-of-day activity and transaction fees. Using ENS address pairs as ground truth (quasi-identifiers), the authors rigorously test their methods and apply their findings to significantly reduce the privacy guarantees of Tornado Cash. \newline
\cite{wu2022tutela} extend on one of \cite{Beres2020} graph representation learning (node-embedding) algorithms and apply it at scale. Further, they also propose a set of new clustering heuristics targeting Tornado Cash, showing that careless user behavior can still reveal identity. They build an application built on those heuristics to measure the anonymity of an Ethereum address. \newline
All of the methods mentioned will be discussed in greater detail in Section 3. \newline
Broadening the scope of entity identification, \cite{victorlüders2019} analyzed the largest Ethereum ERC-20 token networks from a graph perspective. Similarly, \cite{casalebrunet2021} apply a similar analysis to various non-fungible token (NFT) ecosystems. Both find similarities to interaction graphs in social networks (hub-and-spoke). Further, \cite{Payette2017} propose a segmentation of Ethereum address space into four distinct behaviour groups sharing similar attributes using k-means clustering.

\cite{yu2023} propose a novel approach for correlation analysis by exploiting network information, rather than treating users as entities and analyzing on-chain data. Although this approach has the potential to avoid the impact of privacy-enhancing technologies, it introduces a new set of limitations and problems. Thus, this approach may be of great interest for further research, particularly when privacy-enhancing techniques become more widely used.

\textbf{Our contribution}\\
Due to Decentraland's blockchain-based architecture, each avatar contains information about a user's Web3 address. 
To obtain these addresses, a separate project collected snapshots of avatar activity over an ten-month period from July 2022 to April 2023. 
In this work, we perform entity identification on this address set containing 473,927 addresses. 
The applicability of existing heuristics is tested, and if feasible, their efficacy in detecting entities using multiple addresses is evaluated. 
The objective is to estimate the number of real-world entities that are represented by these addresses. Our focus is on clustering addresses within this set alone, while disregarding possible clusters outside of it that were not active in Decentraland during the given timeframe. 
Furthermore, we introduce our own heuristics for identifying address clusters and evaluate their suitability for implementation in other contexts.
%Anwenden/discuss and evaluate existing heuristics to my specific address set.
%Anpassen, erweitern
%Eigene, spezifische Heuristiken ausdenken

The remainder of this paper is structured as follows: In Section 2, we provide a brief overview of the key concepts (basics, necessary for understanding the setup), including Ethereum accounts and the EVM, tokens, Decentraland and privacy-enhancing protocols. In Section 3, we describe our data collection methodology and the data preparation. Section 4  describes different clustering heuristics in detail. Section 5 applies the clustering heuristics from 4. Finally we discuss the results of our analysis and summarize our findings in Sections 5 and 6.

% Motivation \& Relevance of the topic \\
% Transaction data publicly available
% Not really good to hide illicit business acitivities.
% First, this was exploited in Bitcoin and Bitcoins UTXO model
% Recently also account-based
%Literarure Review: Still nascent field, First start with bitcoin/UTXO based, go over to account-based: Victor, Beres et al., Tutela \\

@@

\textbf{Contract Accounts and EOAs}\\
Account-based blockchains usually distinguish between \textit{externally owned accounts} (EOAs) and \textit{contract accounts} (smart contracts). EOAs are created and controlled by private keys and can be used to hold the native protocol asset, send and receive transactions, and interact with contract accounts. Contract accounts are controlled by the contract's code, their state can be modified through transactions sent to the contract and they cannot initiate transactions. \citep{buterin2014ethereum} \newline Each account has a 20-byte address encoded in hexadecimal. For EOAs, this address is based on the last 20 bytes of the Keccak-256 hash of the ECDSA public key. For contract accounts, it is defined as the last 20 bytes of the Keccak-256 hash of the RLP encoding of the sender address and account nonce. \citep{GW:14} \newline
In addition, \cite{eip1014} introduces a new contract creation mechanism allowing for more predictable contract addresses.%EIP-1014

Ethereum Mainnet and Polygon PoS are both account-model blockchains and use the same execution logic, the Ethereum Virtual machine.

- Polygon PoS is a so-called sidechain, closely associated with Ethereum mainnet and used for scaling

%\textbf{Ethereum, Polygon, EVM}\\

\textbf{Transactions and Transfer Events}\\

\textbf{Tokens and Token Standards}\\
%shorter than in Seminar Thesis, comparable to Metaverse Retailing Paper

\textbf{Metaverse and Decentraland} \\
- Much of the Decentraland infrastructure has been deployed on Ethereum Mainnet

\textbf{Privacy-enhancing protocols} \\

>>>>

%The most prominent/widely used non-custodial mixer is Tornado cash (Schär, Tutela).
%If someone obtains information that allows them to link a blockchain address to an entity, they may effectively observe that entity’s entire transaction history and associated activity. Even if the entity uses multiple addresses, any link between these addresses may expose the fact that they belong to the same person.

\textbf{Related work}\\
Previous work in this field can be broadly distinguished between heuristics developed for the unspent transaction output (UTXO) model (e.g, Bitcoin and ZCash) and the account-based model (e.g., Ethereum, Polygon). While both models share the concept of addresses, the notion of accounts is not present in UTXO-based blockchains. The way in which transactions are processed is fundamentally different, and therefore, clustering heuristics are not applicable to both paradigms.

A number of address clustering heuristics in the UTXO setup have been proposed for Bitcoin and derivatives \citep{Androulaki2013, Meiklejohn2013, Haslhofer2016, jourdan2018, kappos2022}. All these methods will not be the focus of this work and therefore not discussed. As suggested by Nakamoto in the Bitcoin whitepaper, most Bitcoin wallet implementations use a new key pair for each transaction to keep them from being linked to a common owner. \newline
In contrast to the UTXO-model, native transactions in account-based blockchains can only move funds between a single sender and receiver, and the ``change``remains in the sender account. Subsequent transactions necessarily use the same address again. The account-based model essentially relies on address reuse on the protocol level. Therefore, privacy guarantees should be lower and most users likely use only a small number of addresses. \citep{Beres2020} \newline
% Make an assumption that people only use a small set of addresses
Clustering heuristics for account-based blockchains were first introduced by \cite{FV:17}. He proposed and applied heuristics that exploit patterns related to deposit addresses, multiple participation in airdrops, and token authorization mechanisms. \newline
\cite{Beres2020} propose more universally applicable methods, as they argue that Victor’s heuristics, while powerful, assume participation in certain on-chain events. Their technique interprets transactions or token transfers as network graphs, with addresses as nodes and asset transfers as edges. They quantitavely compare graph-representation learning algorithms (a subset of machine learning) and propose further user profiling techniques based on time-of-day activity and transaction fees. Using ENS address pairs as ground truth, the authors rigorously test their methods and apply their findings to significantly reduce the privacy guarantees of Tornado Cash. \newline
\cite{wu2022tutela} extend on one of \cite{Beres2020} graph representation learning (node-embedding) algorithms and apply it at scale. Further, they also propose a set of new clustering heuristics targeting Tornado Cash, showing that careless user behavior can still reveal identity. They build an application built on those heuristics to measure the anonymity of an Ethereum address. All of the methods mentioned will be discussed in greater detail in Section 3. \newline
Broadening the scope of entity identification, \cite{victorlüders2019} study the largest Ethereum ERC-20 token networks from a graph perspective. Similarly, \cite{casalebrunet2021} apply a similar analysis to various non-fungible token (NFT) ecosystems. Both find that many of the networks follow either a star or a hub-and-spoke pattern. These patterns are similar to interaction graphs in social networks. Further, \cite{Payette2017} propose a segmentation of Ethereum address space into four distinct behaviour groups sharing similar attributes using k-means clustering. \newline
Rather than treating users as entities and analyzing on-chain data, \cite{yu2023} propose a novel approach for correlation analysis by exploiting network information. Although this approach has the potential to avoid the impact of privacy-enhancing technologies, it introduces a new set of limitations and problems. Thus, this approach may be of great interest for further research, particularly when privacy-enhancing techniques become more widely used.

\textbf{Our contribution}\\
In this work, we perform entity identification on an address set containing 473,927 addresses. To obtain these addresses, a separate project collected snapshots of avatar activity in Decentraland over an ten-month period from July 2022 to April 2023. Due to Decentraland's blockchain-based architecture, each avatar contains information about a user's Web3 address. 
We test the applicability of existing heuristics, and if feasible, their efficacy (in detecting entities using multiple addresses) is evaluated.
The main objective is to estimate the number of real-world entities that these addresses represent. The clustering of addresses is limited to this set only, without considering any (clusters for) addresses that were not registered/active in Decentraland during the given time frame.
Furthermore, we introduce our own heuristics for identifying address clusters and evaluate their suitability for implementation in other contexts.
%Anwenden/discuss and evaluate existing heuristics to my specific address set.
%Anpassen, erweitern

The remainder of this paper is structured as follows: In Section 2, we provide a brief overview of the basic concepts necessary to understand the setup. The include Ethereum accounts and the EVM, tokens, Decentraland, and privacy-enhancing protocols. Section 3 describes our methodology for data collection and preparation. The fourth section provides a detailed explanation of various clustering heuristics. Section 5 applies the clustering heuristics from 4. The fifth section applies the clustering heuristics described in section 4, after which we discuss the results of our analysis and summarize our findings in sections 5 and 6.

@@

\textbf{Contract Accounts and EOAs}\\
Account-based blockchains usually distinguish between \textit{externally owned accounts} (EOAs) and \textit{contract accounts} (smart contracts). EOAs are created and controlled by private keys and can be used to hold the native protocol asset, send and receive transactions, and interact with contract accounts. Contract accounts are controlled by the contract's code, their state can be modified through transactions sent to the contract and they cannot initiate transactions. \citep{buterin2014ethereum} \newline Each account has a 20-byte address encoded in hexadecimal. For EOAs, this address is based on the last 20 bytes of the Keccak-256 hash of the ECDSA public key. For contract accounts, it is usually the last 20 bytes of the Keccak-256 hash of the RLP encoding of the sender address and account nonce. \citep{GW:14}

\textbf{Ethereum, Polygon and the EVM}\\
Ethereum and Polygon are both smart contract based account-model blockchains. They share the same execution logic - the Ethereum Virtual machine (EVM).
The EVM provides a standardized framework for contract execution, ensuring that smart contracts produce deterministic results across all nodes in the network \citep{GW:14}. The EVM is independent of the underlying blockchain protocol. This allows other blockchain implementations to adopt this standardized framework for contract execution.\newline
Polygon is a so-called sidechain and aims to address the scalability limitations of Ethereum. The adoption of the EVM means that both blockchains share the same user address schemes. It further enables developers to deploy and execute Ethereum-based smart contracts on the Polygon network with minimal changes. \citep{matic_whitepaper}
%A high-level programming language is typically used to write smart contracts, and a compiler to convert them into bytecode \citep{mastering_ethereum}. Solidity is currently the dominant smart contract language.
%Polygon PoS is a so-called sidechain, closely associated with Ethereum mainnet and used for scaling.
%Ethereum Mainnet is the chain on which much of the Decentraland infrastructure has been deployed.
%In addition, \cite{eip1014} introduces a new contract creation mechanism allowing for more predictable contract addresses.%EIP-1014
%This allows developers to calculate the address of a contract before it is deployed.

\textbf{Tokens and Token Standards}\\
According to \cite{roth2019tokenization}, \textit{tokens} as rivalrous, digital units of value that represent ownership of an asset or utility. Smart contracts are the primary method of creating tokens on EVM-based blockchains. In essence, a smart-contract-based token is a mapping of accounts with token balances and a set of functions defining how these balances can be changed. Any smart contract containing these elements can be interpreted as a token contract. \citep{roth2019tokenization} \newline
Token standards specify basic interfaces that allow for interoperability between contracts. These standards do not prescribe an implementation but set minimum requirements without restricting the design beyond that. 
The most common token standards are ERC-20\footnote{https://eips.ethereum.org/EIPS/eip-20} for fungible tokens (NFTs), ERC-721\footnote{https://eips.ethereum.org/EIPS/eip-721} for non-fungible tokens or ERC-1155\footnote{https://eips.ethereum.org/EIPS/eip-1155} for semi-fungible tokens.\newline
When a (blockchain) transaction completes, it produces a transaction receipt that contains log entries providing information about the actions that occurred during its execution. The standarization of tokens allows us to listen for ``transfer events``, which are emitted whenever an asset/token changes ownership.\cite{mastering_ethereum}
These transfer events include the \textit{sender} and \textit{recipient} address, along with a \textit{token ID} and/or \textit{amount}.

%Token standards specify basic interfaces that allow for interoperability between contracts. These standards do not prescribe an implementation but set minimum requirements without restricting the design beyond that. ERC-20 (fungible), ERC-721 (non-fungible) or ERC-1155 (semi-fungible) are the most widely used token standards. \newline

%Solidity high-level objects called \textit{Events} construct these logs, which can be queried from a full node and are stored separately from the state.

%Token Definition
%Implemented in a standardized way
%Three most popular token standards
%Transfer emits Events/Logs/Transaction receipts, we can listen for Events

\textbf{Decentraland} \\

%The term metaverse refers to Neal Stephenson’s 1992 novel Snow Crash and describes an immersive virtual world that is populated by humans-as-avatars.
Decentraland is the first large-scale blockchain-based virtual world. Users can own various digital assets, e.g., monetary units, land parcels, avatar apparel, or user names. The ownership of these assets is tracked and managed by smart contracts on Ethereum and Polygon. \citep{goldbergschaer}

 Why interesting? Business perspective

LAND / MANA

%If you choose to experience Decentraland Explorer without a wallet, the information will only be locally stored: you will be able to walk around, customize your Avatar and chat with others in-world, but you won’t have the chance to receive daily rewards, participate in events or log in with a different device using the same Guest ID and Avatar.

\textbf{Non-custodial crypto asset mixers} \\

-> intuition of how it works,  distinction between custodial and non-custodial, Tornado cash = SC based

According to \cite{nadler2023tornado}, crypto-asset mixers are currently the most widely used approach to achieve privacy on public blockchains. On a high level, we can distinguish between two types of implementation.

The simplest version is a custodial model. In this setup, users send their assets to a centralized service provider operating the pool. In return, the provider sends back the assets to a privately relayed recipient address. This approach is fully trust-based because the service provider has complete control over the assets and access to the identifying data.

, where a centralized service provider operates the mixing service. Users send their funds to the mixing service’s public deposit address and specify a recipient address over a private channel.

In contrast to that, non-custodial crypto asset mixers do not require trust. Instead, they rely on cryptographic schemes that allow anyone to prove and independently verify the validity of a withdrawal without disclosing the link to a specific deposit.

Moreover, the user does not have to share the identifying information with anyone and there is no liquidity risk, since the funds are locked and cannot be used in any other way. As such, non-custodial mixers can be created as an immutable and independent infrastructure, where no centralized entity can unilaterally control, alter, or delete the information.

Various entities deposit the same amount of a specific crypto asset to a mixer address. The mixer acts as a pool. Anyone who has contributed to the pool may then generate a new address and withdraw their funds without revealing the link between the depositor and withdrawal addresses. Third parties can still observe the addresses that have deposited to and withdrawn from the pool, but given a large enough anonymity set, those third parties cannot link a specific depositor address to a specific withdrawal address.

The best known smart-contract based non-custodial mixer is Tornado Cash.

%Crypto asset mixers can be implemented at various levels of technological sophistication.
